---
title: "Data Manipulation"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)


```

This document will show how to *manipulate* data. 

Import the two datasets that we're going to manipulate


```{r}
litters_df = read_csv("data/FAS_litters.csv", na = c("NA", "", ".'"))
litters_df = janitor::clean_names(litters_df)

pups_df = read_csv("data/FAS_pups.csv", na = c("NA", "", ".'"))
pups_df= janitor::clean_names(pups_df)  
```

## `select`

Use `select()` to select variables 
```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, group:gd18_weight)

select(litters_df, -pups_survive)


select(litters_df, -(group:gd18_weight))

select(litters_df, starts_with("gd"))

select(litters_df, contains("pups"))

select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)

select(litters_df, litter_number, group)

relocate(litters_df, litter_number, pups_survive)

```
Learning Assesment: In the pups data, select the columns containing litter number, sex, and PD ears.
```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter`

```{r}
filter(litters_df, gd_of_birth == 20)

filter(litters_df, pups_born_alive >= 5)

filter(litters_df, pups_survive != 4)

filter(litters_df, group == "Con7")

filter(litters_df, group %in%c("Low7", "Low8"))

filter(litters_df, group %in%c("Low7", "Low8"), pups_born_alive ==8)

drop_na(litters_df)
drop_na(litters_df, gd0_weight)

```
Learning assesment: In the pups data:
Filter to include only pups with sex 1
Filter to include only pups with PD walk less than 11 and sex 2

```{r}
filter(pups_df, sex == 1)
filter(pups_df, sex == 2, pd_walk < 11)
```

## `mutate`
To select columns or to change them or create new ones.
```{r}
mutate(litters_df, wt_gain = gd18_weight - gd0_weight)

mutate(litters_df, sq_pups = pups_born_alive^2)

mutate(litters_df, group = str_to_lower(group))
```
Learning Assesment:In the pups data:
Create a variable that subtracts 7 from PD pivot
Create a variable that is the sum of all the PD variables
```{r}
mutate(pups_df, pivot_minus7 = pd_pivot - 7)

mutate(pups_df, pd_sum = pd_ears + pd_eyes + pd_pivot + pd_walk)
```


## `arrange`

```{r}

arrange(litters_df,pups_born_alive, gd0_weight)
head(arrange(litters_df, group, pups_born_alive), 10)


```

## PIPING!!!!    


```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |> 
  janitor::clean_names() |>
  select(-pups_born_alive) |>
  filter(group == "Con7") |>
  mutate(
    wt_gain = gd18_weight-gd0_weight
  )
```

