---
title: "Data Import"
output: github_document
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl) 
library(haven)

```

This document will show how to import data. 

## Import the FAS litters CSV

```{r}
litters_df = read_csv("data/FAS_litters.csv")



litters_df = janitor::clean_names(litters_df)

litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval=FALSE}
view(litters_df)
```


Use relative paths 
```{r}
pups_df = read_csv("data/FAS_pups.csv")
 

pups_df= janitor::clean_names(pups_df) 

head(pups_df)
```

Use an absolute path. Does not work once data wrangling folder was moved to bonus_directory. Relative path will still work tho. 
```{r}
pups_df = read.csv("/Users/tinavarela/Desktop/data_wrangling_1/data/FAS_pups.csv")
```

## look at read csv options 

col_names and skipping rows
```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    col_names= FALSE,
    skip = 2
  )
litters_df
```

What about missing data 
```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", ".")
  )
```

What if we code `group` as a factor variable 
```{r}
litters_df = 
  read_csv(
    file = "data/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )
```


## Import an Excel File 

Import MLB 2011 summary data, only use sheet = when there are multiple sheets. This one does not, but is included for practice.  

```{r}
mlb_df = read_excel("data/mlb11.xlsx", sheet = "mlb11")
```

## Import SAS data 
```{r}
pulse_df = read_sas("data/public_pulse_data.sas7bdat")
```

## Never use read.csv, and never use the $ 